\papertitle

### 2020-01-27 paper/paper.tex
Some changes worth of committing

%\documentclass[a4paper, twocolumn]{article}
\documentclass[a4paper]{article}
\newcommand{\papertitle}{Potential of IO-Aware Workflows in Climate and Weather}

@@

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true,urlcolor=red]{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{cleveref}
\usepackage{todonotes}
\usepackage{comment}

@@

{ \color{black}{

\section*{Abstract}
The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other, and, moreover, independently from the needs of end-to-end user workflows that ultimately lead to insight.
As climate and weather workflows become increasingly complex and blend beyond data centres while at the same time, storage hierarchies become deeper, the community investigates means to re-organise storage access to utilise such environments fully.

The key contributions of this paper are:
1) we sketch the visions of an integrated data-driven approach and discuss the challenges and implications of this strategy.
2) we illustrate an architecture and roadmap that allows the seamless integration into existing workflows.
The tools employed here to achieve this extended workflow are Cylc, XIOS, DDN IME, and ESDM.

@@

High-Performance Computing harnesses the fastest hardware components to enable the execution of tightly coupled applications from science and industry.
Typical use-cases cover the numerical simulation of physical systems or the analysis of large-scale observational data.
In the domain of climate and weather, there is a huge demand for the orchestration of ensembles of simulation models and the generation of data products.
A weather forecast service such as ECMWF generates .... \jk{Cylc team, please provide some facts}
Likewise ensembles in climate can ...

Based on their needs, the community has developed a software ecosystem that supports them to execute their large-scale workflows.
While a big leap forward, many processes still require experts.
For example, porting a workflow from one system to another requires to adjust runtime parameters of applications and to decide how data is managed.
Since performance is of key importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target supercomputer.
To obtain best performance, a specific supercomputer typically requires substantial changes to the workflow that tailor the workflow to the specific machine.
For instance, a data-driven workflow may benefit from using a heterogeneous set of computing and storage technology at the same time.

Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
As the complexity of systems increases and various alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimize IO in workflows.

In many cases, modifications are not performed because: 1) they are labor intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) users are not aware of the potential of the complex system.

In this paper, we show that knowledge about workflow execution and overall experimental design helps to
optimise execution of climate and weather workflows and may increase the performance, throughput and cost-efficiency of the environments, providing an incentive to users and data-centres that cannot be neglected.

The paper is organized as follows:
In \Cref{sec:workflows}, we provide an overview for the specification of workflows in climate and weather, and the involved software stack.

@@

In this section, we describe the environment workflows are executed in and the typical software stack.

@@

To satisfy the needs for different workflows, data centers are providing a infrastructure consisting of compute and storage devices with different characteristics -- that make them more efficient for specific tasks.
Take for example the supercomputer Mistral at DKRZ that consists of about 3,000 nodes.
It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
Each node provides a SSD for local storage.
DKRZ has additionally two shared Lustre file systems with different performance characteristics.
The user can access both file systems as a work and scratch file system.
While data is kept on the work file system indefinitely its available space is limited by a quota.
The scratch file system allows to store more data but data is automatically deleted after a period of time.

Future centers are likely to have even more heterogeneity: A variety of accelerators (GPU, TPU, FPGAs), active storage, in-memory, and in-network computing technologies provide storage and processing capabilities.
\Cref{fig:heterogeneous} shows such an environment with a focus on computation and storage.
Some of these technologies might be local to specific compute nodes or globally available.
Depending on the need, the characteristics of storage range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
\jk{Mention BBs, e.g., DDN IME}
Naturally, the variety of tasks executed by a single workflow may benefit from the different storage and computing infrastructure.

@@

The relevant software stack involved in executing workflows is depicted in \Cref{XXX}
\jk{insert picture about software stack here}
To solve a computational problem, we can consider describing the necessary steps from the start to the insight as a workflow.
Cylc executes a workflow description comprising of the individual data processing steps and the dependencies among them.

\subsection{IO Stack}

A typical IO stack for a single parallel application like a climate model is shown in \Cref{fig:layers}.

\begin{minipage}{0.2\textwidth}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{layers-xios}
  \caption{IO Path for an MPI-Parallel Application}
  \label{fig:layers}
\end{figure}
\end{minipage}
\qquad
\begin{minipage}{0.7\textwidth}
In our example, we assume the application is parallelized for performance reasons and uses MPI to coordinate the application.
It may use XIOS to gather data from individual fields which then uses NetCDF4 to store the data as a file.
XIOS provides a domain-level semantics and ...\jk{TODO}
Under the hood, NetCDF4 used the HDF5 API and file format.
Internally, HDF5 uses MPI again and its data types to specify the type of the data stored.
Finally, data is stored on a parallel file system like Lustre, which -- on the server-side -- stores data in a local file system that is stored on block devices and storage media such as SSDs and HDDs.

Different applications involved in a workflow may use a different IO stack to store their outputs.
Naturally, the application that has to input previously generated data must use a compatible API to be able to read the specific data format.
In this figure, for example, XIOS is beneficial for parallel I/O and a process might directly rely on the NetCDF API to read in previous data.
\end{minipage}

@@

Cylc is in charge to execute and monitor workflows of which each step is submitted to the batch scheduler of a data center.
 %, and the relevance of heterogeneous infrastructure to run workflows.
%We focus on high-level considerations to illustrate mandatory concepts.

Consider the Cylc suite configuration for a toy monthly cycling workflow in Figure \ref{fig:cylc}.
In this workflow, a warm-cycled atmospheric model (\textbf{model}) simulates the physics from a current state to predict the future after a timestep, for example, after six hours.
This process is repeated in the model to simulate month and years into the future.
Once the simulation of a month is computed, the data of this month can be analyzed.
A user has specified a detailed workflow in a Cylc suite file (shown on the left).
In this workflow, the \textbf{model} is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks.
Also, consider, in addition, a task \textbf{check} that compares some verification metric against products from two cycles earlier.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\columnwidth]{cylc}
  \caption{Example workflow\cite{TODO}}
  \label{fig:cylc}
\end{figure}

This suite presents the dependencies among the tasks, but it is missing the information about I/O and storage.
The developers make the configuration about parallelism and where the data be stored in a script file that is executed for the specific task.
The decisions about storage must take into consideration the architecture in which the workflow will run.

@@

The scripts for the tasks in Cylc define how a dataset is stored on the available storage systems.
\Cref{fig:lifecycle} shows three possible life cycles for a specific dataset and the access pattern to the dataset.
Consider that the machine has three file systems: a fast scratch file system on which data may reside only for a week, a slower work file system, and each node has a local file system.
The dataset could be stored on the local storage first to avoid congestion on the work file system, then be migrated to work file system where subsequent tasks of the workflow may read it multiple times.
At the end, this dataset might be an intermediate product that can then be deleted.
Alternatively, it could be stored on the scratch file system immediately and accessed there.
However, that would require that the last read access is within the 7-day period before files on scratch are automatically removed.
The last scenario shows the case where data is created on work by a task but then copied to a local node to allow read accesses of subsequent tasks.
This might be beneficial for small random accesses.
However, it would require that subsequent tasks are placed on the same machine where data is now copied to.

The scientist responsible for the workflow must optimise the available resources intuitively.
These are just examples and scientists perform such kind of optimizations already manually.

@@

Other tools for workflow handling, how is this information transported?
dataspaces
ADIOS

\section{Vision for IO-Aware Workflows}

Applied scientists should not spend time understanding hardware characteristics, but using their time to develop their work, let the computer do the job the simplest way possible, and just collect and analyse the results.
However, nowadays, to be able to run a job in a HPC environment efficiently, researchers have to have profound knowledge about their workflow, which is expected, but also about decisions regarding storage, communication, and computing.

We propose an approach to reduce the burden on researchers and, at the same time, optimise the decisions about jobs running in HPC systems.
Once we have an automated decision about where the job will run and how the storage will be managed, the scientists can then run their workflows in any machine without further interactions and even without previous knowledge about the machine architecture.

@@

So far, Cylc allowed to specify tasks, and the dependencies among them.
We aim to enrich this information with characteristics for input, and the products.
An example workflow containing data products is illustrated in \Cref{fig:workflow}.
Arrows indicate dependencies between nodes that represent tasks and data.
In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
Most of the workflow can run automatically except the manual quality control of the products and the final data usage of Product\,3 in the end.
The last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
In this approach, each task is annotated with the required inputs and the generated products.
Each data product is annotated with metadata such as data life cycle information, e.g., the value of data and how long should it be kept.

@@

{ \color{black}{

@@

Now, comes the question. Assuming the information is available, who will be responsible for making the decisions for optimising the storage and the processing tasks? The answer here is simple: the ESDM Scheduler will do it. The ESDM Scheduler can be used in different steps of the process. ESDM has already an interface with NetCDF. Assuming the user can provide extra information in advance, ESDM can interact in the process before the user give the data to the Cylc tool, using the information to construct an optimised workflow and, once Cylc builds the workflow, ESDM can interact again after the configuration of the scripts for both effective and efficient run of the parallel applications regarding the available storage and architecture.

}}

{ \color{black}{

@@

Figure \ref{fig:cycle-io-dep} now introduces one example of dependencies among the datasets and tasks: datasets colours indicate which task will use their information and datasets in yellow represent information that will be used in a different cycle. The tricky part when we consider only tasks and not the datasets in the workflow is that it may lead to the false impression that the tasks only depend on the immediately previous one. Note, for instance, that one dataset produced by Task \textbf{model} is later used in Task \textbf{prod\,gen}. That is not an immediate conclusion when we analyse just the arrows for the dependencies. It is clear that Task \textbf{model} is a dependency for Task \textbf{post\,prod}, which is then a dependency for Task \textbf{prod\,gen}, but the crucial information here is that any dataset currently being used in the current cycle can also be used later. From the figure we can also see that Task \textbf{model} will generate one dataset for Task \textbf{post\,proc}, one dataset for Task \textbf{prod\,gen}, two datasets for Task \textbf{verification}, and three datasets for a later cycle.
The access to this information before the process starts can be used to optimise how these datasets will be stored during the workflow. This information today is static and manual, leaving opportunities for automatisation and optimisations open.
The idea here is to embrace the concept that tasks dependencies are really datasets dependencies. Once we can insert the datasets explicitly in the workflow, it is easier to see that many optimisations regarding storage can be made.

@@

}}

@@

Now:
\begin{enumerate}
  \item User runs Cylc to start a suite
  \item Cylc parses the suite file, generate dependencies and generate a schedule for the execution
  \item Once a task can be executed because dependencies are ready: Cylc generates on the fly a Slurm script with the required metadata for SLURM and invoking the user-defined script for that task and then runs Slurm to queue the script
  \item Slurm queues the task and once the task is ready to be dispatched it is started on the supercomputer
  \item The shell-script executes the users-provided instructions, some of those are to generate file names using Cylc and the variables that are generated by Cylc. These filenames use templates provide by the user that define the storage location to be used.
  \item User-specified applications run taking the filenames as instructed.
  The application may use NetCDF or XIOS to perform the IO.

@@

\begin{enumerate}
  \item User runs Cylc to start a suite.
  \item Cylc parses the suite file, generate dependencies and generate a schedule for the execution.

  Now, Cylc will call into EWIOScheduler to potentially reprioritize the order of execution of tasks.
  The EWIOScheduler will read the suite file and the suite-io.rc file to make suggestions but won't enforce anything.

@@

  Potentially, Cylc will change the metadata of the script telling Slurm to place a job on the same nodes as a previous job to reusue local data.
  Potentially, Cylc may also add as metadata information to migrate data.

@@

  \item Slurm queues the task and once the task is ready to be dispatched it is started on the supercomputer

  If a migration has to be performed before or after a job is run, it would do so.

  \item The shell-script executes the users-provided instructions, some of those are to generate file names using Cylc and the variables that are generated by Cylc.
  These filenames use templates provide by the user that define the storage location to be used.

  The way filenames are generated is now slightly different, using a new command that then uses ESWIOSchedule.
  The guy reads the suite file(s) and generates information for ESDM where to prioritize to place or access data. The filename hence encoded additional information.
  Potentially, we generate an esdm.conf file to reflect the specific hardware we want to use.

@@

    to perform the IO.

  ESDM loads the system configuration knowing the potential targets and to schedule application-local IO.
  Application uses ESDM with the script filename, ESDM extracts the long-term schedule information and considers it during "application-local" IO scheduling.

@@

%, and data products are generated which are processes that are executed outside of the first simulation.

@@

This problem is even more severe for data products: a storage system provides a single level of performance and resilience and performance for stored files.
Theoretically, different fault-tolerance and performance policies can be realised with different storage systems and using management policies.
However, these policies would be coarse-grained and would not account for the priority of the workflow and, most importantly, the value of data.
The importance of data depends on aspects like costs to reproduce the data (can it be reproduced easily), the type of the experiment (test run, production run) and runtime constraints for the overall and potential crucial workflow.
While value and priority should influence fault-tolerance strategies and imply the quality of service for performance and availability, in practice, this is not the case as the storage just sees blocks and files.

@@

>>>>

%\documentclass[a4paper, twocolumn]{article}
\documentclass[a4paper]{article}
\newcommand{\papertitle}{Potential of I/O-Aware Workflows in Climate and Weather}

@@

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true,urlcolor=red]{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{cleveref}
\usepackage{todonotes}
\usepackage{comment}
\usepackage{ulem}

@@

\section*{Abstract}
The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
\sout{{\color{blue}{
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other, and, moreover, independently from the needs of end-to-end user workflows that ultimately lead to insight.}}}
{\color{cyan}{
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other and from the needs of end-to-end user.}}
As climate and weather workflows become increasingly complex and blend beyond data centres while, at the same time, storage hierarchies become deeper, the community investigates means to re-organise storage access to utilise such environments fully.

The key contributions of this paper are:
1) we sketch the visions of an integrated data-driven approach and discuss the challenges and implications of this strategy.
2) we illustrate architecture and roadmap that allows the seamless integration into existing workflows.
The tools employed here to achieve this extended workflow are Cylc, XIOS, DDN IME, and ESDM.

@@

High-Performance Computing (HPC) harnesses the fastest hardware components to enable the execution of tightly coupled applications from science and industry.
Typical use-cases cover the numerical simulation of physical systems and the analysis of large-scale observational data.
In the domain of climate and weather, there is a considerable demand for the orchestration of ensembles of simulation models and the generation of data products.
A weather forecast service such as ECMWF generates .... \jk{Cylc team, please provide some facts}
Likewise, ensembles in climate can ...

Based on their needs, the {\color{cyan}{HPC}} community has developed a software ecosystem that supports them to execute their large-scale workflows.
While the current advances correspond to a big leap forward, many processes still require experts.
For example, porting a workflow from one system to another requires to adjust runtime parameters of applications and to decide how data is managed.
Since performance is of crucial importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target supercomputer.
To obtain the best performance, a specific supercomputer typically requires substantial changes to the workflow to tailor the workflow to the particular machine.
For instance, a data-driven workflow may benefit from using a heterogeneous set of computing and storage technology at the same time.

Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
As the complexity of systems increases and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise I/O in workflows. {\color{cyan}{First time I/O is mentioned. I would change to something like... it becomes increasingly difficult even for experts to manually optimise resources (other stuff) and, in particular, I/O in such workflows.}}
In many cases, modifications are not performed because: 1) they are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) users are not aware of the potential of the complex system.

In this paper, we show that knowledge about workflow execution and overall experimental design helps to
optimise the execution of climate and weather workflows and may increase the performance, throughput and cost-efficiency of the environments, providing an incentive to users and data-centres that cannot be neglected.

The paper is organised as follows:
In \Cref{sec:workflows}, we provide an overview of the specification of workflows in climate and weather and the involved software stack.

@@

In this section, we describe how the environment workflows are executed in the typical software stack.

@@

To satisfy the needs for different workflows, data centres are providing a infrastructure consisting of computing and storage devices with different characteristics which makes them more efficient for specific tasks.
Take, for example, the supercomputer Mistral at DKRZ that consists of about {\color{cyan}{about?}} 3,000 nodes.
It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
Each node provides an SSD for local storage and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
The user can access both file systems \sout{{\color{blue}{as a work and scratch file system}}}. {\color{cyan}{as work and scratch file systems.}}
While data is kept on the work file system indefinitely, its available space is limited by a quota.
The scratch file system allows to store more data, but data is automatically deleted after some time.

Future centres are expected to have even more heterogeneity. A variety of accelerators (GPU, TPU, FPGAs), active storage, in-memory, and in-network computing technologies provide storage and processing capabilities.
\Cref{fig:heterogeneous} shows such an environment with a focus on computation and storage.
Some of these technologies might be local to specific compute nodes or globally available.
Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
\jk{Mention BBs, e.g., DDN IME}
Naturally, the variety of tasks executed by a single workflow may benefit from the different storage and computing infrastructure.

@@

The relevant software stack involved in executing workflows is depicted in \Cref{XXX}
\jk{insert picture about software stack here}
To solve a computational problem, we can consider describing the necessary steps from the start to the insight {\color{cyan}{insight? I really don't undestand some of your uses of this word.}} as a workflow.
Cylc executes a workflow description comprising of the individual data processing steps and the dependencies among them.

\subsection{I/O Stack}

A typical I/O stack for a single parallel application like a climate model is shown in \Cref{fig:layers}.

\begin{minipage}{0.2\textwidth}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{layers-xios}
  \caption{I/O Path for an MPI-Parallel Application}
  \label{fig:layers}
\end{figure}
\end{minipage}
\qquad
\begin{minipage}{0.7\textwidth}
In our example, we assume the application is parallelised for performance reasons and uses MPI to \sout{{\color{blue}{coordinate}}} {\color{cyan}{run}} the application.
It may use XIOS to gather data from individual fields which then uses NetCDF4 to store the data as a file.
XIOS provides a domain-level semantics and ...\jk{TODO}
Under the hood, NetCDF4 used the HDF5 API and file format.
Internally, HDF5 uses MPI again and its data types to specify the nature of the data stored.
\sout{{\color{blue}{Finally, data is stored on a parallel file system like Lustre, which, on the server-side, stores data in a local file system that is stored on block devices and storage media such as SSDs and HDDs.}}}
{\color{cyan}{Finally, data is stored on a parallel file system like Lustre, which, on the server-side, stores data in local file system as block devices on storage media such as SSDs and HDDs.}}

Different applications involved in a workflow may use a different I/O stack to store their outputs.
Naturally, the application that inputs previously generated data must use a compatible API to be able to read the specific data format.
In this figure, for example, XIOS is beneficial for parallel I/O and a process might directly rely on the NetCDF API to read previous data.
\end{minipage}

@@

Cylc is in charge to execute and monitor workflows in which each step is submitted to the batch scheduler of a data centre.
 %, and the relevance of heterogeneous infrastructure to run workflows.
%We focus on high-level considerations to illustrate mandatory concepts.

Consider the Cylc suite configuration for a toy monthly cycling workflow in Figure \ref{fig:cylc}.
In this workflow, a warm-cycled atmospheric model (\textbf{model}) simulates the physics from a current state to predict the future after a timestep, for example, after six hours.
This process is repeated in the model to simulate month and years into the future.
Once the simulation of a month is computed, the data for this month can be analysed.
{\color{cyan}{This is wuite strange for me. It means that only after a month is completed one will use the information to start previsions? I would expect it to be more like a window-month than a fixed month, no?}}
A user has specified a detailed workflow in a Cylc suite file (shown on the left).
In this workflow, the \textbf{model} is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks.
Also, consider, in addition, a task \textbf{check} that compares some verification metric against products from two cycles earlier.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\columnwidth]{cylc}
  \caption{Example workflow\cite{8675433}}
  \label{fig:cylc}
\end{figure}

This suite presents the dependencies among the tasks, but it is missing the information about I/O and storage.
Decisions about storage must take into consideration the architecture in which the workflow will run.
The developers make the configuration about parallelism and where the data is stored in a script file that is executed for each specific task.

@@

The scripts for the tasks in Cylc define how the data is stored on the available storage systems.
\Cref{fig:lifecycle} shows three possible life cycles and the access pattern for a specific dataset.
Consider that the machine has three file systems: a fast \textbf{scratch} file system on which data may reside only for a week, a slower \textbf{work} file system, and each node has a \textbf{local} file system.
The dataset could be stored on the \textbf{local} storage first to avoid congestion on the \textbf{work} file system, then be migrated to \textbf{work} file system where subsequent tasks of the workflow may read it multiple times.
In the end, this dataset might be an intermediate product that can then be deleted.
Alternatively, it could be stored on the \textbf{scratch} file system immediately and accessed there.
However, that would require that the last read access happens before files on \textbf{scratch} are automatically removed.
The last scenario shows the case where data is created on \textbf{work} by a task, but then it is copied to a \textbf{local} node to allow read accesses of subsequent tasks which might be beneficial for small random accesses.
However, it would require that the subsequent tasks are placed on the same machine where data is now stored.
The scientist responsible for the workflow must optimise the available resources intuitively.
These are just examples and currently the scientists perform such kind of optimisations manually.

@@

Other tools for workflow handling, how is this information transported?

dataspaces

ADIOS

\section{Vision for I/O-Aware Workflows}

Applied scientists should not spend time understanding hardware characteristics, but using their time to develop their work, let the computer do the job the simplest way as possible, and just collect and analyse the results.
However, nowadays, to be able to run a job in a HPC environment efficiently, researchers have to have profound knowledge about their workflow, which is expected, but also about decisions regarding storage, communication, and computing.

We propose an approach to reduce the burden on researchers and, at the same time, optimise the decisions about jobs running in HPC systems.
Once we have an automated decision about where the job will run and how the storage will be managed, scientists can then run their workflows in any machine without further interactions and even without previous knowledge about the machine architecture.

@@

So far, Cylc allowed specifying tasks and the dependencies among them.
We aim to enrich this information with characteristics for input, and the products. {\color{cyan}{What?!}}
An example workflow containing data products is illustrated in \Cref{fig:workflow}.
\sout{{\color{blue}{Arrows indicate dependencies between nodes that represent tasks and data.}}}
{\color{cyan}{Nodes represent tasks and data and arrows indicate dependencies among tasks.}}
In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
In this approach, each task is annotated with the required inputs and the generated products.
Each data product is annotated with metadata such as data life cycle information, e.g., the value of data and how long should it be kept.

@@

@@

Now, comes the question. Assuming the information is available, who will be responsible for making the decisions for optimising the storage and the processing tasks? The answer here is simple: the ESDM Scheduler will do it. The ESDM Scheduler can be used in different steps of the process. ESDM has already an interface with NetCDF. Assuming the user can provide extra information in advance, ESDM can interact in the process before the user give the data to the Cylc tool, using the information to construct an optimised workflow. Once Cylc builds the workflow, ESDM can interact again after the configuration of the scripts for both effective and efficient run of the parallel applications regarding the available storage and architecture.

@@

Figure \ref{fig:cycle-io-dep} now introduces one example of dependencies among the datasets and tasks: datasets colours indicate which task will use their information and datasets in yellow represent information that will be used in a different cycle. The tricky part when we consider only tasks and not the datasets in the workflow is that it may lead to the false impression that the tasks only depend on the immediately previous one. Note, for instance, that one dataset produced by Task \textbf{model} is later used in Task \textbf{prod\,gen}. That is not an immediate conclusion when we analyse just the arrows for the dependencies. It is clear that Task \textbf{model} is a dependency for Task \textbf{post\,prod}, which is then a dependency for Task \textbf{prod\,gen}. Still, the crucial information here is that any dataset currently being used in the current cycle can also be used later. From the figure we can also see that Task \textbf{model} will generate one dataset for Task \textbf{post\,proc}, one dataset for Task \textbf{prod\,gen}, two datasets for Task \textbf{verification}, and three datasets for a later cycle.
The access to this information before the process starts can be used to optimise how these datasets will be stored during the workflow. This information today is static and manual, leaving opportunities for automatisation and optimisations open.
The idea here is to embrace the concept that tasks dependencies are really datasets dependencies. Once we can insert the datasets explicitly in the workflow, it is easier to see that many optimisations regarding storage can be made.

@@

@@

Now:
\begin{enumerate}
  \item user runs Cylc to start a suite
  \item Cylc parses the suite file, generate dependencies and generate a schedule for the execution
  \item Once a task can be executed because dependencies are ready: Cylc generates on the fly a Slurm script with the required metadata for SLURM and invoking the user-defined script for that task and then runs Slurm to queue the script
  \item Slurm queues the task, and once the task is ready to be dispatched it is started on the supercomputer
  \item The shell-script executes the users-provided instructions, some of those are to generate file names using Cylc and the variables that are generated by Cylc. These filenames use templates provided by the user that defines the storage location to be used.
  \item User-specified applications run taking the filenames as instructed.
  The application may use NetCDF or XIOS to perform the I/O.

@@

\begin{enumerate}
  \item user runs Cylc to start a suite.
  \item Cylc parses the suite file, generate dependencies and generate a schedule for the execution.

  Now, Cylc will call into EWIOScheduler to potentially reprioritise the order of execution of tasks.
  The EWIOScheduler will read the suite file and the suite-io.rc file to make suggestions but won't enforce anything.

@@

  Potentially, Cylc will change the metadata of the script telling Slurm to place a job on the same nodes as a previous job to reuse local data.
  Potentially, Cylc may also add as metadata information to migrate data.

@@

  \item Slurm queues the task, and once the task is ready to be dispatched it is started on the supercomputer

  If migration has to be performed before or after a job is run, it would do so.

  \item The shell-script executes the users-provided instructions, some of those are to generate file names using Cylc and the variables that are generated by Cylc.
  These filenames use templates provided by the user that defines the storage location to be used.

  The way filenames are generated now slightly different, using a new command that then uses ESWIOSchedule.
  The guy reads the suite file(s) and generates information for ESDM where to prioritise to place or access data. The filename hence encoded additional information.
  Potentially, we generate an esdm.conf file to reflect the specific hardware we want to use.

@@

    to perform the I/O.

Knowing the potential targets, ESDM loads the system configuration to schedule application-local I/O.
  The application uses ESDM with the script filename, ESDM extracts the long-term schedule information and considers it during "application-local" I/O scheduling.

@@

%, and data products are generated, which are processes that are executed outside of the first simulation.

@@

This problem is even more severe for data products: a storage system provides a single level of performance and resilience and performance for stored files.
Theoretically, different fault-tolerance and performance policies can be realised with varying storage systems and using management policies.
However, these policies would be coarse-grained and would not account for the priority of the workflow and, most importantly, the value of data.
The importance of data depends on aspects like costs to reproduce the data (can it be reproduced easily), the type of the experiment (test run, production run) and runtime constraints for the overall and potential crucial workflow.
While value and priority should influence fault-tolerance strategies and imply the quality of service for performance and availability, in practice, this is not the case as the storage just sees blocks and files.

@@

\bibliographystyle{alpha}
\bibliography{paper}
